---
title: "Homework, Week 4"
output: html_notebook
---

```{r}
library(rethinking)

set.seed(100)
```

# Q1

```{r}
# Compute entropy
H <- function(p) -sum(p * log(p))

# Compute K-L divergence
D_KL <- function(p, q) sum(p * (log(p) - log(q)))
```

```{r}
islands = list()

islands[[1]] = c(0.2, 0.2, 0.2, 0.2, 0.2)
islands[[2]] = c(0.8 ,0.1, 0.05, 0.025, 0.025)
islands[[3]] = c(0.05, 0.15, 0.7, 0.05, 0.05)

sapply(islands, H)
```

Island 1 has the largest entropy as it is most uncertain, with the most even distribution. Island 2 has the smallest entropy, as its distribution is uneven.

```{r}
DM <- matrix(NA, nrow=3, ncol=3)

# rows: model; cols: true
for (j in 1:3)
  for (i in 1:3)
    DM[i, j] <- D_KL(islands[[j]], islands[[i]])

round(DM, 2)
```

The first model has the smallest K-L divergence with the other two models, as it is most uncertain and therefore least surprised by the others.

# Q2.

```{r}
d <- sim_happiness(seed=1977, N_years=1000)

precis(d)
```

```{r}
d2 <- d[d$age>17,] # only adults
d2$A <- (d2$age - 18) / (65 - 18)
```

```{r}
d2$mid <- d2$married + 1

m6.9 <- quap(
  alist(
    happiness ~ dnorm(mu, sigma),
    mu <- a[mid] + bA*A,
    a[mid] ~ dnorm(0, 1),
    bA ~ dnorm(0, 2),
    sigma ~ dexp(1)
  ), data=d2)

precis(m6.9, depth=2)
```

```{r}
m6.10 <- quap(
  alist(
    happiness ~ dnorm(mu, sigma),
    mu <- a + bA*A,
    a ~ dnorm(0, 1),
    bA ~ dnorm(0, 2),
    sigma ~ dexp(1)
  ), data=d2)

precis(m6.10)
```

```{r}
compare(m6.9, m6.10)
```

The model (m6.9) which includes marriage status as a predictor is expected to make better predictions, as it has a lower WAIC; while marriage status is a collider, it does provide associational information when conditioned on (included in the model).

However, as marriage status is a collider, we should not simply choose the model with lower WAIC if we are interested in a causal interpretation: m6.10 shows no causal effect of age on happiness.