---
title: "Homework, Week 5"
output: html_notebook
---

```{r}
library(rethinking)

set.seed(100)
```

# Q1

```{r}
# load the data
data(Wines2012)
d <- Wines2012

# standardize score and get index variables
dat_list <- list(
  S = scale(d$score),
  judge_id = as.integer(d$judge),
  wine_id = as.integer(d$wine)
)
```

```{r}
dens(dat_list$S)
```

```{r}
# define the model formulae
flist <- alist(
  S ~ dnorm(mu, sigma),
  mu <- bJ[judge_id] + bW[wine_id],
  bJ[judge_id] ~ dnorm(0, 1),
  bW[wine_id] ~ dnorm(0, 1),
  sigma ~ dexp(1)
)

set_ulam_cmdstan(TRUE)

# fit the model
model <- ulam(flist, data=dat_list, chains=4, cores=4)
```

```{r}
precis(model, depth=2)

traceplot(model)
```

```{r}
precis_plot(precis(model, depth=2))
```

There is more variation among judges than among wines.

# Q2

```{r}
# standardize score and get index variables
dat_list <- list(
  S = scale(d$score),
  F = ifelse(d$flight == "red", 1, 2),
  JA = d$judge.amer + 1,
  WA = d$wine.amer + 1
)
```

* `JA` is 1 for French and 2 for American judges.
* `WA` is 1 for French and 2 for New Jersey wines.
* `F` is 1 for red and 2 for white wines.

## Indicator variable model

```{r}
flist <- alist(
  S ~ dnorm(mu, sigma),
  mu <- bF*F + bJA*JA + bWA*WA,
  c(bF, bJA, bWA) ~ dnorm(0, 0.5),
  sigma ~ dexp(1)
)

model <- ulam(flist, data=dat_list, chains=4, cores=4)
```

```{r}
precis_plot(precis(model, depth=2), xlim=c(-1.25, 1.25))

precis(model)
```

The `bF` intercept seems to show that red and white wines are scored about the same. American judges were more positive, but American wines were judged more negatively.

## Index variable model

```{r}
flist <- alist(
  S ~ dnorm(mu, sigma),
  mu <- bF[F] + bJA[JA] + bWA[WA],
  bF[F] ~ dnorm(0, 0.5),
  bJA[JA] ~ dnorm(0, 0.5),
  bWA[WA] ~ dnorm(0, 0.5),
  sigma ~ dexp(1)
)

model <- ulam(flist, data=dat_list, chains=4, cores=4)
```

```{r}
precis_plot(precis(model, depth=2), xlim=c(-1.25, 1.25))

precis(model, depth=2)
```

The interpretation of these results is consistent with the indicator variable model. To show that the models are the same, we can do:

```{r}
post <- extract.samples(model)
diff_wa <- post$bWA[,2] - post$bWA[,1]
precis(diff_wa)
```

This equals the value of bWA in the indicator variable model.

# Q3

