---
title: "Homework, Week 2"
output: html_notebook
---

```{r}
library(rethinking)

set.seed(100)
```

# Q1

We will build a model to predict height for the !Kung adults, using weight as a predictor.

## Load the !Kung data

```{r}
# load the data
data(Howell1)
d <- Howell1

# show the data frame 
print(d)

# summarize the data
precis(d)
```
## Filter for the adults

```{r}
# filter for the adults
d2 <- d[ d$age >= 18 , ]

# show the data frame
print(d2)

# summarize the data
precis(d2)
```

```{r}
# plot height distribution
dens(d2$height)

# plot height versus weight
plot(d2$height ~ d2$weight)
```

## Constructing a model

We will model height using a Normal likelihood with unknown mean and unknown variance. The unknown mean can be modelled with an intercept and a slope for the height predictor.

A reasonable prior for the intercept could be a Normal distribution centred on 178cm (the author's height), with 95% of the probability between $178 \pm 40$, meaning a standard deviation of 20. As we have seen, and it makes sense, that height tends to increase with weight, it seems reasonable to choose a prior that restricts the slope to positive values, such as a $\text{Log-Normal}(0, 1)$.

The standard deviation must be positive, but unlike the mean we do not really have a natural intuition for plausible values, so we can use a flat prior that is bounded between (say) 0 and 50. A standard deviation of 50 would imply that 95% of individual heights lie within 100cm of the average height, which is a large range.

```{r}
# plot prior for intercept mean
curve(dnorm(x, 178, 20), from=100, to=250)

# plot prior for intercept standard deviation
curve(dunif(x, 0, 50), from=-10, to=60)

# plot prior for slope 
curve(dlnorm(x, 0, 1), from=-5, to=10)
```

```{r}
# prior predictive simulation

N <- 100
a <- rnorm(N, 178, 20) # intercept
b <- rlnorm(N, 0, 1) # slope

plot(NULL, xlim=range(d2$weight), ylim=c(-100, 400),
  xlab="weight", ylab="height")
abline(h=0, lty=2)
abline(h=272, lty=1, lwd=0.5)
mtext("log(b) ~ dlnorm(0, 1)")
weight_bar <- mean(d2$weight)
for (i in 1:N)
  curve(a[i] + b[i]*(x - weight_bar),
    from=min(d2$weight), to=max(d2$weight), add=TRUE,
    col=col.alpha("black", 0.2)
  )
```

```{r}
# define the model formulae
flist <- alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * (weight - mean(d2$weight)),
    a ~ dnorm(178, 20),
    b ~ dlnorm(0, 1),
    sigma ~ dunif(0, 50)
  )

# fit model using quadratic approximation
model <- quap(flist, data=d2)

precis(model)
```

```{r}
# plot posterior inference against the data

plot(height ~ weight, data=d2, col=rangi2)

post <- extract.samples(model)
a_map <- mean(post$a)
b_map <- mean(post$b)
curve(a_map + b_map * (x - mean(d2$weight)), add=TRUE)
```

## Individual predictions

```{r}
post <- extract.samples(model)

mu_at_45 <- post$a + post$b * (45 - mean(d2$weight))

dens(mu_at_45 , col=rangi2, lwd=2, xlab="mu|weight=45")

PI(mu_at_45)
```

```{r}
q1 <- data.frame(weight=c(45, 40, 65, 31))
height_sim <- sim(model, data=q1)

expected_height <- apply(height_sim, 2, mean)
ci_height <- apply(height_sim, 2, PI)

q1$expected_height <- expected_height
q1$L89 <- ci_height[1,]
q1$U89 <- ci_height[2,]
round(q1, 1)
```

# Q2

We will model the relationship between height (cm) and the natural log of weight (log-kg) for all ages.

```{r}
d$log_weight <- log(d$weight)

plot(height ~ log_weight, data=d)
```

```{r}
# define the model formulae
flist <- alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * (log_weight - mean(d$log_weight)),
    a ~ dnorm(178, 20),
    b ~ dlnorm(0, 1),
    sigma ~ dunif(0, 50)
  )

# fit model using quadratic approximation
model <- quap(flist, data=d)

precis(model)
```

```{r}
# plot posterior inference against the data

plot(height ~ weight, data=d, col=rangi2)

log_weight_seq <- log(1:60)

mu <- sim(model, data=list(log_weight=log_weight_seq))

mu_mean <- apply(mu, 2, mean)
mu_ci <- apply(mu, 2, PI)

lines(exp(log_weight_seq), mu_mean, lwd=3)

shade(mu_ci, exp(log_weight_seq))
```