---
title: "Homework, Week 6"
output: html_notebook
---

```{r}
library(rethinking)
library(dagitty)

set.seed(100)
```

# Q1

```{r}
# load the data
data(NWOGrants)
d <- NWOGrants

d
```

```{r}
# draw DAG of the setup
g <- dagitty('dag {
    G [pos="0,1"]
    D [pos="1,0"]
    A [pos="2,1"]
    
    G -> D -> A
    G -> A
}')

plot(g)
```

To get the total causal effect of `gender` on `awards`:

$A_i \sim \text{Binomial}(N_i, p_i)$ \
$\text{logit}(p_i) = \alpha_{\text{GID}[i]}$
$\alpha_j \sim \text{Normal}(0, 1.5)$

```{r}
# get index variables
dat_list <- list(
  awards = as.integer(d$awards),
  apps = as.integer(d$applications),
  gender = ifelse(d$gender == "m", 1, 2)
)

dat_list
```

* `gender` is 1 for male and 2 for female applicants.

```{r}
# define the model formulae
flist <- alist(
  awards ~ dbinom(apps, p),
  logit(p) <- a[gender],
  a[gender] ~ dnorm(0, 1.5)
)

# fit the model
model <- ulam(flist, data=dat_list, chains=4, cores=4)
```

```{r}
precis_plot(precis(model, depth=2))

precis(model, depth=2)
```

Male applicants seem to have a higher award rate, but by how much?

```{r}
post <- extract.samples(model)
diff <- inv_logit(post$a[,1]) - inv_logit(post$a[,2])
precis(list(diff=diff))
```

The rate for male applicants is about 3% higher than for females on the absolute scale.

To get the direct causal effect of `gender` on `awards`, we can condition on department:

$A_i \sim \text{Binomial}(N_i, p_i)$ \
$\text{logit}(p_i) = \alpha_{\text{GID}[i]} + \delta_{\text{DEPT}[i]}$
$\alpha_j \sim \text{Normal}(0, 1.5)$
$\delta_k \sim \text{Normal}(0, 1.5)$

```{r}
# get index variables
dat_list <- list(
  awards = as.integer(d$awards),
  apps = as.integer(d$applications),
  disc = as.integer(d$discipline),
  gender = ifelse(d$gender == "m", 1, 2)
)

dat_list
```

* `gender` is 1 for male and 2 for female applicants.

```{r}
# define the model formulae
flist <- alist(
  awards ~ dbinom(apps, p),
  logit(p) <- a[gender] + b[disc],
  a[gender] ~ dnorm(0, 1.5),
  b[disc] ~ dnorm(0, 1.5)
)

# fit the model
model <- ulam(flist, data=dat_list, chains=4, cores=4)
```

```{r}
precis_plot(precis(model, depth=2))

precis(model, depth=2)
```

```{r}
post <- extract.samples(model)
diff <- post$a[,1] - post$a[,2]
precis(list(diff=diff))
```
On the relative scale, the award rate for male applicants is 14% higher than for females.

```{r}
postcheck(model)
```

# Q2

```{r}
# draw DAG of the setup
g <- dagitty('dag {
    G [pos="0,1"]
    D [pos="1,0"]
    A [pos="2,1"]
    U [pos="3,0"]
    
    G -> D -> A
    G -> A
    U -> D
    U -> A
}')

plot(g)
```
Here, U is an unobserved variable. Conditioning on department now opens up a backdoor path from D to A through U, and we cannot get an unconfounded estimate of the effect of gender on awards.